_Paenibacillus_ species resolution for 16S rRNA V12 and V4 regions.

### 2018-04-01 inst/manuscript/manuscript_supplemental.Rmd

## Background
16S rRNA amplicon sequencing is commonly used for microbial community characterization, including differential abundance analysis. 
A limitation to 16S rRNA amplicon sequencing is a lack of taxonomic resolution, where organisms are only identifiable to the genus or family level. 
We define taxonomic resolution as the ability to differentiate between groups within a taxonomic level, for example differentiating between species within a genus. 
Taxonomic resolution can vary by clade and amplicon regions, though the extent to which it varies is not well characterized.

Here we demonstrate how `metagenomeFeatures` and the `MgDb` annotation packages can be used to characterize taxonomic resolution for a specific clade and amplicon region, specifically for the _Paenibacillus_ genus and V12 and V4 regions 
Using the Greengenes 13.5, `greengenes13.5MgDb` we will first subset the database to obtain _Paenibacillus_ taxonomic and sequence data. 
Next we will explore the _Paenibacillus_ species composition, then characterize the within and between species sequence similarity for the 16S rRNA gene V12 and V4 regions.

## Required Packages  
Our analysis uses the `DECIPHER` package to extract the amplicon regions, perform multiple sequence alignment, and generate a pairwise sequence distance matrix. 
The `tidyverse` and `ggpubr` packages will be used to reformat the taxonomic and distance matrix data and generate summary figures.

@@

## _Paenibacillus_ Sequence and Taxonomy Data
We first subset the Greeengenes 13.5 database using the `mgDb_select` function. 
Then summarize the taxonomy data using the `tidyverse` package. 
<!-- Loading metagenomeFeatues and greengenes db -->

```{r}
paeni_16S <- mgDb_select(gg13.5MgDb, 
                          type = c("taxa","seq"), 
                          keys = "Paenibacillus", 
                          keytype = "Genus")
```

@@

Generating pairwise distance matrix for taxonomic resolution analysis and converting distance matrix to data frame for analysis.

>>>>

## Background
16S rRNA amplicon sequencing is commonly used for microbial community characterization, including differential abundance analysis. 
A limitation to 16S rRNA amplicon sequencing is a lack of taxonomic resolution, where organisms are only identifiable to the genus or family level. 
We define taxonomic resolution as the ability to differentiate between groups within a taxonomic level, for example differentiating between species within a genus. 
While similar to determining whether a sequence represents a novel species, here we are only interested in determining whether the 16S rRNA region of interest contains sufficient information for species level taxonomic assignment.
Taxonomic resolution can vary by clade and amplicon regions, though the extent to which it varies has not been characterized.

Here we demonstrate how `metagenomeFeatures` and `MgDb` annotation packages can be used to characterize taxonomic resolution for a specific clade and amplicon region, specifically for the _Paenibacillus_ genus and V12 and V4 regions 
Using the Greengenes 13.5, `greengenes13.5MgDb` we will first subset the database to obtain _Paenibacillus_ taxonomic and sequence data. 
Next we will explore the _Paenibacillus_ species composition, then characterize the within and between species sequence similarity for the 16S rRNA gene V12 and V4 regions.

## Required Packages  
Our analysis uses the `DECIPHER` package to extract the amplicon regions, perform multiple sequence alignment, and generate a pairwise sequence distance matrix [@decipher]. 
The `tidyverse` and `ggpubr` packages will be used to reformat the taxonomic and distance matrix data and generate summary figures [@tidyverse;@ggpubr].

@@

## _Paenibacillus_ Sequence and Taxonomy Data
We first subset the Greeengenes 13.5 database using the `mgDb_select` function. 
Then summarize the taxonomy data using functions from `tidyverse` package, specifically `dplyr`, `stringr` and `forcats` functions for manipulating `data.frames`, `strings`, and `factor` vectors respectively. 
<!-- Loading metagenomeFeatues and greengenes db -->

```{r}
paeni_16S <- metagenomeFeatures::mgDb_select(gg13.5MgDb, 
                          type = c("taxa","seq"), 
                          keys = "Paenibacillus", 
                          keytype = "Genus")
```

@@

Generating pairwise distance matrix using the `DistanceMatrix` function in the `DECIPHER` package for taxonomic resolution analysis and converting distance matrix to data frame for analysis.
